Paradigms in Physics 2.0

### 2017-06-29 paper.tex

\begin{abstract}
In 2016, the Department of Physics at Oregon State University began a
process to revise our Paradigms in Physics curriculum for physics
majors.  We began with a
survey of students and faculty, followed by individual interviews with
the faculty teaching each existing course.  As we developed a plan to
address student- and faculty-identified challenges in the curriculum,
we met with each faculty member individually to explain and refine our
proposal, which was unanimously approved by the faculty.  Major
changes include major changes to several courses (math
methods, computational physics, modern physics, electronics, and
classical mechanics), including the introduction of two sophomore-year
courses designed specifically to help prepare students for their
upper-division courses.
\end{abstract}

\maketitle
% * <ethan.minot@oregonstate.edu> 2017-06-28T19:21:10.109Z:
%
% ^.
%\section{Introduction}
%\davidsays{Why does a reader care? A dramatic example of curriculum change.
%A dramatic revision of said.  So others can adopt this method of change.  This is
%an implementation of education research.  It is an exemplar for the community.
%}
The Paradigms in Physics project began in 1996, when three faculty
members at Oregon State applied to the NSF for funding to redesign the
upper-division curriculum.  The motivation for the original Paradigms~1.0
project was in fact similar to the motivation for Paradigms~2.0: a
desire to soften the ``brick wall'' encountered by students upon
reaching the junior year, within the constraint that transfer students
must be able to graduate with two years of upper-division
courses~\cite{manogue2001paradigms}.  On top of this, there was a
desire to provide students with a broad knowledge of physics prior to
the GRE exams in the fall of their senior year.  The process of
redesigning the curriculum was dominated by the creation of index cards
listing subject content, and sorting these cards into courses.  This
process involved the entire faculty, and culminated in a unanimous
agreement to adopt the resulting curriculum.

@@

After incorporating the feedback from  individual meetings with every
faculty member, we scheduled
two full faculty meetings with a week in between.  In the first meeting,
we presented our proposal in detail and invited questions, but not discussion.
This was needed for a couple of reasons:  those faculty we met with first
may not have seen the final proposal, and some faculty during their
individual meetings chose to focus on a small subset of the
curriculum, often relating to courses they had themselves taught.
During the following week faculty engaged in hallway discussions of the proposal.
This week helped ensure that the faculty did not feel rushed into a vote.
In the second faculty meeting, we again presented our proposal, and opened
the floor for discussion.  After considerable discussion, largely on
changes that were not part of the proposal, we unanimously voted to adopt
the proposed changes.

%\newcommand\mathcourse[2]{\emph{#1 (#2)}}
\newcommand\mathcourse[2]{\emph{#1}}
\newcommand\noted[2]{\textbf{#1} (#2)}
\newcommand\paradigm[1]{{\sc #1} (3)}
\newcommand\capstone[1]{#1 (3)}
\newcommand\onecredit[1]{#1 (1)}
\newcommand\threecredit[1]{#1 (3)}
\newcommand\fourcredit[1]{#1 (4)}

@@

\subsection{5-week \emph{Paradigm} courses}
To begin with the most distinctive feature of our curriculum, we
wanted to maintain the existing intensive 7-hour-per-week schedule in
the junior year, which we have found effective.  This schedule allows students to focus
intensely on a single topic, it fosters the building of a cohort of students, and 
the daily schedule is helpful for active engagement.  We chose to
change from three 3-week courses per quarter to two 5-week courses
per quarter.  This gives students who fall behind a chance to catch up, and enables
us to give more feedback to students prior to the final exam.  Having 3-credit courses
simplifies the setting of teaching loads, and by reducing the number of courses, 
we allow the curriculum to be taught with fewer faculty, albeit at an increased work

>>>>

\begin{abstract}
In 2016, the Department of Physics at Oregon State University began a
process to revise our Paradigms in Physics curriculum for physics
majors.  We began with a colloquium to inform the department of our
plans and request their assistance, followed by a
survey of students and faculty as well as individual interviews with
the faculty teaching each course.  As we developed a plan to
address student- and faculty-identified challenges in the curriculum,
we met with each faculty member individually to explain and refine our
proposal, which was unanimously approved by the faculty.  Major
changes include major changes to several courses (math
methods, computational physics, modern physics, electronics, and
classical mechanics), including the introduction of two sophomore-year
courses designed specifically to help prepare students for their
upper-division courses.
\end{abstract}

\maketitle
% * <ethan.minot@oregonstate.edu> 2017-06-28T19:21:10.109Z:
%
% ^.
%\section{Introduction}
%\davidsays{Why does a reader care? A dramatic example of curriculum change.
%A dramatic revision of said.  So others can adopt this method of change.  This is
%an implementation of education research.  It is an exemplar for the community.
%}
The Paradigms in Physics project began in 1996, when three faculty
members at Oregon State applied to the NSF for funding to redesign the
upper-division curriculum.  The motivation for the original Paradigms~1.0
project was in fact similar to the motivation for Paradigms~2.0: a
desire to soften the ``brick wall'' encountered by students upon
reaching the junior year, within the constraint that transfer students
must be able to graduate with two years of upper-division
courses~\cite{manogue2001paradigms}.  On top of this, there was a
desire to provide students with a broad knowledge of physics prior to
the GRE exams in the fall of their senior year.  The process of
redesigning the curriculum was guided by the creation of index cards
listing subject content, and sorting these cards into courses.  This
process involved the entire faculty, and culminated in a unanimous
agreement to adopt the resulting curriculum.

@@

After incorporating the feedback from  individual meetings with every
faculty member, we scheduled
two full faculty meetings with a week in between.  In the first meeting,
we presented our proposal in detail and invited questions, but not discussion.
This was needed for a couple of reasons:  those faculty we met with first
may not have seen the final proposal, and some faculty during their
individual meetings chose to focus on a small subset of the
curriculum, often relating to courses they had themselves taught.
During the following week faculty engaged in hallway discussions of the proposal.
This week helped ensure that the faculty did not feel rushed into a vote.
In the second faculty meeting, we again presented our proposal, and opened
the floor for discussion.  After considerable discussion, largely on
changes that were not part of the proposal, the faculty unanimously voted to adopt
the proposed changes.

@@

\subsection{5-week \emph{Paradigm} courses}
To begin with the most distinctive feature of our curriculum, we
wanted to maintain the existing intensive 7-hour-per-week schedule in
the junior year, which we have found effective.  This schedule allows students to focus
intensely on a single topic, it fosters the building of a cohort of students, and 
the daily schedule is helpful for active engagement.  We chose to
change from three 3-week courses per quarter to two 5-week courses
per quarter.  This gives students who fall behind a chance to catch up, and enables
us to give more feedback to students prior to the final exam.  Having 3-credit courses
simplifies the setting of teaching loads, and by reducing the number of courses, 
we allow the curriculum to be taught with fewer faculty, albeit at an increased
work load per professor.
